### 2019-03-14 reports/thesis/chapters/5_inference.tex
Write up empirical null

@@

@@

@@

@@

It was chosen that convergence was met when either 10 update steps were taken or when
\begin{equation}
  \left|
    \dfrac{
      \partial
    }
    {
      \partial z
    }
  \ln\widehat{p}_Z(z)
  \right|
  <10^{-5}
\end{equation}
at the current step. This was chosen arbitrary to speed up the algorithm without losing too much accuracy. At the end of the algorithm, for a successful convergence it was required, in addition, that
\begin{equation}
  \left.
    \dfrac{
      \partial^2
    }
    {
      \partial z^2
    }
    \ln\widehat{p}_Z(z)
  \right|_{z=\widehat{\mu}_0}
  < 0 \ .
\end{equation}
Following a successful convergence, the estimator $\widehat{\sigma}_0$ was calculated straight away.

The algorithm does depend on the initial value so using different initial values is recommended. It was chosen here to use the quartiles of the data as the initial values and then select the best solution, the one with the largest $\ln\widehat{p}_Z\left(\widehat{\mu}_0\right)$, out of all the different initial values. Should the algorithm still fail, a random data point is selected to be the initial value until a successful convergence is achieved. The Newton-Raphson method can fail if it converges to a stationary point other than a global maximum, for example a local maximum or a point of inflection.

\subsection{Approximate Properties of $\widehat{\sigma}_0$}

@@

The choice of the bandwidth $h$ is important for the kernel density estimator. It controls how smooth the kernel density estimator is, with higher values of $h$ producing smoother curves \citep{friedman2001elements}. Cross validation methods do exist \citep{bowman1984alternative, sheather2004density} but they can be computationally expensive. Rules of thumb \citep{silverman1986density, sheather2004density} can be used instead and are typically of the form
\begin{equation}
  h = bn^{-1/5}\times\text{min}\left(s_z,\text{IQR}_z/1.34\right)
\end{equation}
where $b=0.9$ \citep{silverman1986density}. This rule of thumb was developed with consideration of bimodal Normal distributions \citep{silverman1986density}. Other options include $b=1.06$ and $b=1.144$ to produce smoother curves \citep{silverman1986density, sheather2004density}.

The kernel density estimator is only used to obtain values of $\widehat{\mu}_0$ and $\widehat{\sigma}_0$. Thus a good bandwidth, in the context of the empirical null, is one which has good properties of $\widehat{\mu}_0$ and $\widehat{\sigma}_0$ rather than the density estimate. However exact properties of these estimators based on the kernel density estimators can be rather complicated. Empirical properties were studied here.

An experiment was conducted to investigate the behaviour of the estimators, $\widehat{\mu}_0$ and $\widehat{\sigma}_0$, on a dataset of $n$ simulated all null $\normal(0,1)$. Various values of $h$ and $n$ were investigated. For a given $h$ and $n$, 100 values of $\widehat{\mu}_0$ and $\widehat{\sigma}_0$ were obtained by repeating the simulation. The median squared error was obtained by taking the median over all squared errors from the 100 estimates. Figure \ref{fig:inference_BandwidthSelection_error} shows the resulting median squared errors.

@@

Overestimates and underestimates are both dangerous and it can affect the false positive and false negative rate. This is because $\widehat{\sigma}_0$ was used to rescale the test statistics accordingly. Recall that the corrected test statistics are
\begin{equation}
  t_i = \dfrac{
    z_i - \widehat{\mu}_0
  }
  {
    \widehat{\sigma}_0
  }
  \ .
  \tag{\ref{eq:inference_zeta}}
\end{equation}
Low values of $\widehat{\sigma}_0$ produce large values in magnitude of $t_i$, potentially testing more of these test statistics as positive. High values of $\widehat{\sigma}_0$ does the opposite. Large number of positives can led to more false positives and vice versa.

The optimal bandwidth for a given $n$ was found numerically. Figure \ref{fig:inference_BandwidthSelection2_plot} shows the log squared error against the bandwidth for various $n$. A smoothing spline \citep{friedman2001elements} was fitted and this was optimised to find the bandwidth which minimises the log squared error for a given $n$.

@@

A relationship between the optimal bandwidth and $n$ was attempted to be found. It was assumed such a relationship has the linear from
\begin{equation}
  h_{\text{optimal}} = b n^{-1/5} + a
\end{equation}
where $b$ and $a$ are parameters to be fitted. Figure \ref{fig:inference_BandwidthSelection2_ruleOfThumb} show the fitting of an identity link gamma GLM in order to estimate $b$ and $a$. The estimated parameters are shown in Table \ref{table:inference_BandwidthSelection2_ruleOfThumb}.

@@

The results showed that the rules of thumb can be improved if a small bias is added to it. This is because an intercept was found when fitting a linear relationship between the optimal bandwidth and $n^{-1/5}$. Adding this bias should improve the performance of the estimator $\widehat{\sigma}_0$. The gradient is similar to the rules of thumb in the literature \citep{sheather2004density}.

@@

A further example is given here. A $200\times200$ sub-sample, containing a defect, was taken as shown in Figure \ref{fig:inference_alt_empirical_z_image}. The histogram of the test statistics, along with the empirical null, is shown in Figure \ref{fig:inference_alt_empirical_z_histogram}. It can be seen that the null distribution is not centred at zero, however the empirical null takes that into account. The estimation of the empirical null parameters are robust as it only depends on the density estimate at the mode only, it should not be affected by non-null test statistics.

\begin{figure}
  \centering
    \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/alt_empirical_z_image.eps}
        \caption{$z$ image}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/alt_empirical_z_image_2.eps}
        \caption{Sample from left}
    \end{subfigure}
    }
    \caption{$z$ image from comparing a scan with \texttt{aRTist}. A sub-sample containing a defect is shown. Highlighted in red are significant pixels.}
    \label{fig:inference_alt_empirical_z_image}
\end{figure}

\begin{figure}
  \centering
    \centerline{
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/alt_empirical_z_histo_null.eps}
        \caption{Histogram and empirical null}
        \label{fig:inference_alt_empirical_z_histogram}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{../figures/inference/alt_empirical_z_p_values.eps}
        \caption{Ordered $p$-values}
        \label{fig:inference_alt_empirical_z_p_values}
    \end{subfigure}
    }
    \caption{Histogram and $p$-values, corrected using the empirical null, of the test statistics taken from the sub-sample.}
\end{figure}

The BH procedure was conducted, corrected for the empirical null. The $p$-values are shown in Figure \ref{fig:inference_alt_empirical_z_p_values}. $\input{../tables/alt_empirical_n_sig.txt}$ pixels were significant. The significant pixels are shown in red in Figure \ref{fig:inference_alt_empirical_z_image}. The majority of the significant pixels are cluster together and highlighting the defect. Not all of the defects were tested positive, but a good number of them are. This should be enough to raise suspicion in that particular area.

Only a few pixels were falsely tested as positive, however they are typically isolated single pixels. Isolated positive pixels should be discarded as they are more than likely to be tested positive by random chance. The technician should of improved the resolution of the x-ray apparatus if one is looking for defects, a or two pixel in size, for the given resolution.

>>>>

\begin{equation*}

@@

\end{equation*}

@@

\begin{multline*}

@@

\end{multline*}

@@

It was chosen that convergence was met when either 10 update steps were taken or when
\begin{equation}
  \log\left|
    \left.
    \dfrac{
      \partial
    }
    {
      \partial z
    }
  \ln\widehat{p}_Z(z)
  \right|_{z=\widehat{\mu}_0}
  \right|
  <-5
\end{equation}
at the current step. This was chosen arbitrary to speed up the algorithm without losing too much accuracy. At the end of the algorithm, for a successful convergence it was required, in addition, that
\begin{equation}
  \left.
    \dfrac{
      \partial^2
    }
    {
      \partial z^2
    }
    \ln\widehat{p}_Z(z)
  \right|_{z=\widehat{\mu}_0}
  < 0 \ .
\end{equation}
Following a successful convergence, the estimator $\widehat{\sigma}_0$ was calculated straight away.

The algorithm does depend on the initial value so using different initial values were used. Further initial values were generated by sampling from $\normal(z^{(0)}, s_z^2)$. This was done multiple times until 3 valid solutions were obtained. The best solution, the one with the largest $\ln\widehat{p}_Z\left(\widehat{\mu}_0\right)$, out of all the different initial values was used as the final answer. The Newton-Raphson method can fail if it converges to a stationary point other than a global maximum, for example a local maximum or a point of inflection.

@@

The bandwidth $h$ controls how smooth the kernel density estimator is, higher values producing smoother curves \citep{friedman2001elements}. Cross validation methods to select $h$ do exist \citep{bowman1984alternative, sheather2004density} but they can be computationally expensive. Rules of thumb \citep{silverman1986density, sheather2004density} can be used instead and are typically of the form
\begin{equation}
  h = bn^{-1/5}\times\text{min}\left(s_z,\text{IQR}_z/1.34\right)
\end{equation}
where $b=0.9$ \citep{silverman1986density}. This rule of thumb was developed with consideration of bimodal Normal distributions \citep{silverman1986density}. Other options include $b=1.06$ and $b=1.144$ to produce smoother curves \citep{silverman1986density, sheather2004density}.

The kernel density estimator was only used to obtain values of $\widehat{\mu}_0$ and $\widehat{\sigma}_0$. Thus a good bandwidth, in the context of the empirical null, is one which has good properties of $\widehat{\mu}_0$ and $\widehat{\sigma}_0$ rather than the density estimate. Exact properties of these estimators based on the kernel density estimators can be rather complicated. For example numerous approximations were needed to show that $\widehat{\sigma}_0$ is an unbiased estimator of $\sigma_0$ to the first order, where higher order terms are functions of $\widehat{\mu}_0 - \mu_0$ and $h$. This is shown in Appendix ?.

An experiment was conducted to investigate how the estimators, $\widehat{\mu}_0$ and $\widehat{\sigma}_0$, varied with $h$ and $n$. This was done on a simulated dataset of $n$ standard Normal random variables. For a given $h$ and $n$, 100 values of $\widehat{\mu}_0$ and $\widehat{\sigma}_0$ were obtained by repeating the simulation of the datasets. The median squared error was obtained by taking the median over all squared errors from the 100 estimates. Figure \ref{fig:inference_BandwidthSelection_error} shows the resulting median squared errors.

@@

Overestimates and underestimates are both dangerous and it can affect the false positive and false negative rate. This is because $\widehat{\sigma}_0$ was used to rescale the test statistics accordingly. Recall that the normalised test statistics are $t_i = \dfrac{z_i - \widehat{\mu}_0}{\widehat{\sigma}_0}$. Low values of $\widehat{\sigma}_0$ produce large values in magnitude of $t_i$, potentially testing more of these test statistics as positive. High values of $\widehat{\sigma}_0$ does the opposite. Large number of positives can led to more false positives and vice versa.

The optimal bandwidth for a given $n$ was found numerically. Figure \ref{fig:inference_BandwidthSelection2_plot} shows the log squared error against the bandwidth for various $n$. A smoothing spline \citep{friedman2001elements} was fitted and this was optimised to find the bandwidth which minimises the log squared error for a given $n$. \texttt{MATLAB}'s \texttt{fit(x,y,'smoothingspline')} function was used to fit the smoothing spline which has inbuilt cross validation procedures. The optimisation was done using \texttt{MATLAB}'s gradient free \texttt{fminsearch} function, using $h=0.9n^{-1/5}$ as the initial value.

@@

A relationship between the optimal bandwidth and $n$ was attempted to be found. It was assumed such a relationship has the linear from
\begin{equation}
  h_{\text{optimal}} = b n^{-1/5} + a
\end{equation}
where $b$ and $a$ are parameters to be fitted. This was chosen to establish a simple relationship with the rules of thumb. Figure \ref{fig:inference_BandwidthSelection2_ruleOfThumb} show the fitting of an identity link Gamma GLM in order to estimate $b$ and $a$. The estimated parameters are shown in Table \ref{table:inference_BandwidthSelection2_ruleOfThumb}.

@@

The rules of thumb can be improved if a small bias is added to it. This is because an intercept was found when fitting a linear relationship between the optimal bandwidth and $n^{-1/5}$. Adding this bias should improve the performance of the estimator $\widehat{\sigma}_0$. The gradient is similar to the rules of thumb in the literature \citep{sheather2004density}.

The GLM fit in Figure \ref{fig:inference_BandwidthSelection2_ruleOfThumb} appeared to underfit, which suggest there may be a more complicated relationship between the optimal bandwidth and $n$. The method used in fitting the smoothing spline could contribute to the underfitting. For example if the smoothing spline was not smooth enough so that local minima appear, it would add variance to the location of the optimal bandwidth. A basic linear relationship was chosen to avoid overfitting and keep the rule of thumb simple.

@@
